---
title: "Reticulate/Python Virtual Environment Startup"
output:
  html_document:
    df_print: paged
---

## Getting Started with _virtualenv_

The first thing to do, when getting started with the `reticulate` package in R, is to decide how you want to manage your python environment. 

In terms of this R project:

- Which version of ython do you want to use?
- Where is this installation of python located?
- What python libraries do you want to make available?

Even if your answer to these questions is: default, default, and I'm not sure yet, I still recommend learning about and using Python virtual environments. There are several ways to create a Python virtual environment. This document describes the process of setting up _virtualenv_ and how to link that environment to `reticulate` in R. 

>_virtualenv_ is a tool to create isolated Python environments. _virtualenv_ creates a folder which contains all the necessary executables to use the packages that a Python project would need. 

[Link: Python Guide for Virtual Environments](https://docs.python-guide.org/dev/virtualenvs/).

I like creating python virtual environments to do work with `reticulate`. There are ways to create virtual environments directly from the reticulate package, but I prefer to create them in the terminal such that the directory and activation script is easily (by default) co-located with the R project directory I'm working in. This is purely a personal workflow preference. 

**Create a virtualenv in the Terminal**

1. Make sure you have python available: `python --version`
2. Make sure you have pip available: `pip --version`
3. Install virtualenv: `pip install virtualenv`
4. Change into the project directory you want to use: `cd my_project_folder`
5. Create a virtualenv: `virtualenv my_project`
6. Activate the virtualenv: `source my_project/bin/activate`
7. Use pip to install python packages (ex. install `requests`): `pip install requests`
8. To exit the virtualenv: `deactivate`

_Clean version of the above commands:_
```
python --version
pip --version
pip install virtualenv
cd my_project_folder
virtualenv my_project
source my_project/bin/activate
pip install requests
```

**Workflow Notes Continued:** When using git in this directory, I do not commit the contents of my virtualenv directory. Instead, I'll use a terminal command to "freeze" the contents of my pip list into a file called `requirements.txt` and check that into version control instead. Another user can then create their own virtualenv by installing directly from that requirements file._

```
pip freeze > requirements.txt
pip install -r requirements.txt
```

## Getting Started in R

1. Start/Restart R
2. Use an existing python virtualenv called `reticulate`, set `required = TRUE`
3. Check the output of `py_config()` to verify that reticulate is set correctly

```{r}
library(reticulate)
use_virtualenv("reticulate", required = TRUE)
py_config()
```


### Environment Module Check:

Compare the output of pip list to make sure you have the expected modules available (_shown in R and python_):

```{r}
# R Code Chunk (uses reticulate)
os <- import("os")
os$system("pip list")
```


```{python}
# Python Code Chunk
import os
os.system("pip list")
```


